# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r load, echo=TRUE}
data<-read.csv("activity.csv", colClasses = c("numeric", "Date", "numeric"))
#deleting NA for preprocessing
data_preprocess<-data[complete.cases(data$steps),]

```


## What is mean total number of steps taken per day?
```{r mean, echo=TRUE}
total<-tapply(data_preprocess$steps,data_preprocess$date,sum);
plot(1:length(total),as.vector(total),type="h",col = "red", lwd = 4,xlab="days",ylab="steps");
mean_steps<-mean(as.vector(total))
median_steps<-median(as.vector(total))
```
### The mean is `r mean_steps` and median is `r median_steps`.

## What is the average daily activity pattern?
```{r daily, echo=TRUE}
average_activity<-tapply(data_preprocess$steps,data_preprocess$interval,mean,simplify=FALSE);
plot(names(average_activity),average_activity,type="l",xlab="interval",ylab="average steps");
```
### The maximum number of steps is with interval of `r names(which.max(average_activity))`

## Imputing missing values
```{r missing,echo=TRUE}
ip_missing<-nrow(data)-sum(complete.cases(data))
data_without_NA<-data
#replace NA's with average daily activity pattern.
df<-data.frame(interval=names(average_activity),steps=average_activity)
count<-0;
for(i in 1:nrow(data_without_NA)){
    if(is.na(data_without_NA$steps[i])){
      for(j in 1:nrow(df)){
        if(df$interval[j]==data_without_NA$interval[i]){
          count=count+1;
          data_without_NA$steps[i]<-as.numeric(df$steps[j]);     
          break;
        }        
      }    
    }
  }
total<-tapply(data_without_NA$steps,data_without_NA$date,sum);
plot(1:length(total),as.vector(total),type="h",col = "green", lwd = 4,xlab="days",ylab="steps");
mean_steps<-mean(as.vector(total))
median_steps<-median(as.vector(total))
```
### We have replaced NA's in "steps" with average daily activity pattern.
### The number of missing values are `r ip_missing`.
### The new mean is `r mean_steps` and median is `r median_steps`.
### Few points to note 
* The mean remains same.
* The median is changed.
* Days in the histogram have increased because of no NA's.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday,echo=TRUE}
library(ggplot2)
index<-grepl("Sunday",weekdays(data_without_NA$date),ignore.case=TRUE) | grepl("Saturday",weekdays(data_without_NA$date),ignore.case=TRUE);
fweek = factor(index,labels=c("weekday","weekend"));
datawithweek<-cbind(data_without_NA,fweek);

X<-aggregate(x = datawithweek[,1], by = list(datawithweek$interval,datawithweek$fweek), FUN = "mean");
names(X)<-c("interval","fweek","steps")
g<-ggplot(X,aes(interval,  steps))+geom_line()
g<-g+geom_point(aes(color = fweek),size = 2, alpha = 1/2)
g<-g+facet_grid(fweek~.) 
g<-g+labs(x = "Interval", y = "Number of steps") 
print(g)
```